 Group Report

### 2024-12-09 draft.Rmd
Add files via upload
this is the draft in markdown

---
title: "R Notebook"
output:
  html_notebook: default
  pdf_document: default
editor_options: 
  markdown: 
    wrap: 72
---

# Group Report

@@

```{r include = FALSE}
echo=FALSE #load all the libraries 
install.packages("leaps") 
library(car) 
library(ggplot2) 
library(reshape2) 
library(tidyverse) 
library(broom) 
library(caret) 
library(RColorBrewer) 
library(leaps) 
required_packages <- c("rmarkdown", "knitr", "evaluate")
install.packages(required_packages)
options(repos = c(CRAN = "https://cran.rstudio.com/"))

@@

```{r echo = FALSE, fig.cap = "Table 1. Preview of the High School Alcoholism and Academic Performance dataset.\n"}
# load the dataset
data <- read.csv("dataset.csv")
head(data, 3)
#cat("Table 1. Preview of the High School Alcoholism and Academic Performance dataset.\n")
```

@@

# Remove rows where either G1 or G2 is equal to 0
data <- subset(data, G1 != 0 & G2 != 0)
head(data)
#cat("Table 2. Preview of the cleaned High School Alcoholism and Academic Performance dataset.\n")
```

```{r echo = FALSE, fig.cap = "Figure 1. Distribution of average of G1 and G2.\n"}
# Distribution histogram of the Average of G1 and G2
hist <- ggplot(data, aes(x = G1_G2_Average)) +
  geom_histogram(binwidth = 1, fill = "skyblue", color = "black", alpha = 0.7) +
  labs(title = "Distribution of average of G1 and G2", x = "Average", y = "Count") +
  theme_minimal()
hist
#cat("Figure 1. Distribution of average of G1 and G2.\n")
```

@@

```{r echo = FALSE, fig.cap="\nFigure 2. Correlation heatmap.\n" }
#setup the covariance matrix
cor_matrix <- cor(data[, sapply(data, is.numeric)], use = "complete.obs")
#print(cor_matrix)
cor_data <- melt(cor_matrix)
ggplot(cor_data, aes(Var1, Var2, fill = value)) +
  geom_tile(color = "white") +
  scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
                       midpoint = 0, limit = c(-1, 1), space = "Lab", 
                       name = "Correlation") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
  coord_fixed() +
  labs(title = "Correlation Heatmap", x = "", y = "")
#cat("\nFigure 2. Correlation heatmap.\n")
```

@@

```{r echo = FALSE, fig.cap = "\nFigure 3. Residual plot for full model.\n"}
# residual plot again fitted values for full model
plot(full_model$fitted.values, residuals(full_model),
     xlab = "Fitted Values",
     ylab = "Residuals",
     main = "Residuals vs Fitted")
abline(h = 0, col = "red", lty = 2)
#cat("\nFigure 3. Residual plot for full model.\n")
```

@@

```{r echo = FALSE, fig.cap = "\nFigure 4. Q-Q plot for full model.\n"}
# QQ-plot for full model
qqnorm(residuals(full_model))
qqline(residuals(full_model), col = "red")
#cat("\nFigure 4. Q-Q plot for full model.\n")
```

@@

```{r echo = FALSE, fig.cap = "\nFigure 5. Q-Q plot for actual model.\n"}
qqnorm(residuals(model))
qqline(residuals(model), col = "red")
#cat("\nFigure 5. Q-Q plot for actual model.\n")
```

@@

```{r echo = FALSE, fig.cap ="\nFigure 6. Residual plot for actual model.\n" }
plot(model$fitted.values, residuals(model), 
     xlab = "Fitted Values", 
     ylab = "Residuals", 
     main = "Residual Plot")
abline(h = 0, col = "red", lty = 2)
#cat("\nFigure 6. Residual plot for actual model.\n")
```

@@

```{r echo = FALSE, fig.cap ="\nFigure 7. Q-Q plot for log-transformed model.\n" }
qqnorm(residuals(model2))
qqline(residuals(model2), col = "red")
#cat("\nFigure 7. Q-Q plot for log-transformed model.\n")
```

```{r echo = FALSE, fig.cap ="\nFigure 8. Residual plot for log-transformed model.\n" }
plot(model2$fitted.values, residuals(model2), 
     xlab = "Fitted Values", 
     ylab = "Residuals", 
     main = "Residual Plot")
abline(h = 0, col = "red", lty = 2)
#cat("\nFigure 8. Residual plot for log-transformed model.\n")
```

@@

#### Coefficient Significance $(α = 0.05)$ / Confidence Intervals $(95\%)$

-   **Intercept**: \$\beta\_0\$ ( \[8.23, 10.9\] )

@@

-   **School**: \$\beta\_1\$ ( \[-2.09, -0.741\] )

    The $p-value$ $5.35 \times 10^{-5}$ is less than $0.05$, so we
    reject $H_0$. This result indicates that the type of school is
    significantly associated with average student performance.

@@

-   **Medu**: \$\beta\_2\$ ( \[-0.0659, 0.496\] )

@@

-   **Higher**: \$\beta\_3\$ ( \[1.33, 3.48\] )

@@

-   **Absences**: \$\beta\_4\$ ( \[-0.199, -0.0629\] )

    The $p-value$ $2.01 \times 10^{-4}$ is less than $0.05$, so we
    reject $H_0$. This indicates that absences are significantly
    associated with lower average student performance.

>>>>

---
title: " Group Report"
output:
  pdf_document: default
  html_notebook: default
editor_options: 
  markdown: 
    wrap: 72
---

@@

```{r include = FALSE}
library(car) 
library(ggplot2) 
library(reshape2) 
library(tidyverse) 
library(broom) 
library(caret) 
library(RColorBrewer) 
library(leaps) 
library(rmarkdown)
library(knitr)
library(evaluate)

@@

```{r include= FALSE}
# load the dataset
data <- read.csv("dataset.csv")
head(data, 1)
cat("Table 1. Preview of the High School Alcoholism and Academic Performance dataset.\n")
```

@@

# Remove rows where either G1 or G2 is equal to 0
data <- subset(data, G1 != 0 & G2 != 0)
head(data,1)
cat("Table 1. Preview of the cleaned High School Alcoholism and Academic Performance dataset.\n")
```

```{r echo = FALSE, fig.cap = "  Distribution of average of G1 and G2.\n", fig.width=4, fig.height=2}
# Distribution histogram of the Average of G1 and G2
hist <- ggplot(data, aes(x = G1_G2_Average)) +
  geom_histogram(binwidth = 1, fill = "skyblue", color = "black", alpha = 0.7) +
  labs(title = "Distribution of average of G1 and G2", x = "Average", y = "Count") +
  theme_minimal()
hist
#cat("Figure 1. Distribution of average of G1 and G2.\n")
```

@@

```{r echo = FALSE, fig.cap="\n. Correlation heatmap.\n", fig.width=4, fig.height=4 }
#setup the covariance matrix
cor_matrix <- cor(data[, sapply(data, is.numeric)], use = "complete.obs")
#print(cor_matrix)
cor_data <- melt(cor_matrix)
ggplot(cor_data, aes(Var1, Var2, fill = value)) +
  geom_tile(color = "white") +
  scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
                       midpoint = 0, limit = c(-1, 1), space = "Lab", 
                       name = "Correlation") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
  coord_fixed() +
  labs(title = "Correlation Heatmap", x = "", y = "")
#cat("\nFigure 2. Correlation heatmap.\n")
```

@@

```{r echo = FALSE, fig.cap = "\n. Residual plot for full model.\n", fig.width=5, fig.height=4}
# residual plot again fitted values for full model
plot(full_model$fitted.values, residuals(full_model),
     xlab = "Fitted Values",
     ylab = "Residuals",
     main = "Residuals vs Fitted")
abline(h = 0, col = "red", lty = 2)
#cat("\nFigure 3. Residual plot for full model.\n")
```

@@

```{r echo = FALSE, fig.cap = "\n. Q-Q plot for full model.\n", fig.width=5, fig.height=4}
# QQ-plot for full model
qqnorm(residuals(full_model))
qqline(residuals(full_model), col = "red")
#cat("\nFigure 4. Q-Q plot for full model.\n")
```

@@

```{r echo = FALSE, fig.cap = "\n. Q-Q plot for actual model.\n", fig.width=5, fig.height=4}
qqnorm(residuals(model))
qqline(residuals(model), col = "red")
#cat("\nFigure 5. Q-Q plot for actual model.\n")
```

@@

```{r echo = FALSE, fig.cap ="\n. Residual plot for actual model.\n", fig.width=5, fig.height=4}
plot(model$fitted.values, residuals(model), 
     xlab = "Fitted Values", 
     ylab = "Residuals", 
     main = "Residual Plot")
abline(h = 0, col = "red", lty = 2)
#cat("\nFigure 6. Residual plot for actual model.\n")
```

@@

```{r echo = FALSE, fig.cap ="\n. Q-Q plot for log-transformed model.\n", fig.width=5, fig.height=4 }
qqnorm(residuals(model2))
qqline(residuals(model2), col = "red")
#cat("\nFigure 7. Q-Q plot for log-transformed model.\n")
```

```{r echo = FALSE, fig.cap ="\n. Residual plot for log-transformed model.\n", fig.width=5, fig.height=4 }
plot(model2$fitted.values, residuals(model2), 
     xlab = "Fitted Values", 
     ylab = "Residuals", 
     main = "Residual Plot")
abline(h = 0, col = "red", lty = 2)
#cat("\nFigure 8. Residual plot for log-transformed model.\n")
```

@@

#### Coefficient Significance ($\alpha$ = 0.05) / Confidence Intervals $(95\%)$

-   **Intercept**: $\beta_0$ ( [8.23, 10.9] )

@@

-   **School**: $\beta_1$ ( [-2.09, -0.741] )

    The $p-value$ $5.35 \times10^{-5}$ is less than $0.05$, so we reject
    $H_0$. This result indicates that the type of school is
    significantly associated with average student performance.

@@

-   **Medu**: $\beta_2$ ( [-0.0659, 0.496] )

@@

-   **Higher**: $\beta_3$ ( [1.33, 3.48] )

@@

-   **Absences**: $\beta_4$ ( [-0.199, -0.0629] )

    The $p-value$ = $2.01\times 10^{-4}$ is less than $0.05$, so we reject
    $H_0$. This indicates that absences are significantly associated
    with lower average student performance.
